\documentclass[submission,copyright,creativecommons]{eptcs}
\providecommand{\event}{NCL'22} % Name of the event you are submitting to
%\usepackage{breakurl}             % Not needed if you use pdflatex only.
\usepackage{underscore}           % Only needed if you use pdflatex.
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{enumerate}
\usepackage{prftree}
%\usepackage{ntheorem}
%\usepackage{tikz-cd}
\usepackage{hyperref}
%\usepackage{float}
%\usepackage{graphicx}
\usepackage{quiver}
\usepackage[all,cmtip]{xy}
\usepackage{bussproofs}
\usepackage{proof}
%\usepackage[a4paper, total={6in, 8in}]{geometry}
%\usepackage{lscape}
%% \theorembodyfont{}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{defn}{Definition}[section]
\newtheorem{example}{Example}[section]
%\newtheorem{defn}[definition]{Definition}
%\newtheorem{example}[definition]{Definition}
%\newtheorem*{proof}{Proof : }
\newtheorem{fact}[theorem]{Fact}
\makeatletter
\newsavebox{\@brx}
\newcommand{\llangle}[1][]{\savebox{\@brx}{\(\m@th{#1\langle}\)}%
  \mathopen{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\newcommand{\rrangle}[1][]{\savebox{\@brx}{\(\m@th{#1\rangle}\)}%
  \mathclose{\copy\@brx\kern-0.5\wd\@brx\usebox{\@brx}}}
\makeatother
\newcommand{\ldbc}{[\![}
\newcommand{\rdbc}{]\!]}
\newcommand{\tbar}{[\vec{x}/\vec{t}]}
\newcommand{\ltbar}{[\vec{x}, x/\vec{t}, x]}
\newcommand{\tl}{\otimes \mathsf{L}}
\newcommand{\tr}{\otimes \mathsf{R}}
\newcommand{\lright}{{\multimap}\mathsf{R}}
\newcommand{\lleft}{{\multimap}\mathsf{L}}
\newcommand{\pass}{\mathsf{pass}}
\newcommand{\unitl}{\mathsf{IL}}
\newcommand{\unitr}{\mathsf{IR}}
\newcommand{\ax}{\mathsf{ax}}
\newcommand{\id}{\mathsf{id}}
\newcommand{\ot}{\otimes}
\newcommand{\lolli}{\multimap}
\newcommand{\illol}{\rotatebox[origin=c]{180}{$\multimap$}}
\newcommand{\I}{\mathsf{I}}
\newcommand{\msfL}{\mathsf{L}}
\newcommand{\defeq}{=_{\mathsf{df}}}
\newcommand{\comp}{\mathsf{comp}}
\newcommand{\RI}{\mathsf{RI}}
\newcommand{\LI}{\mathsf{LI}}
\newcommand{\Pass}{\mathsf{P}}
\newcommand{\F}{\mathsf{F}}
\newcommand{\xvdash}{\vdash^{x}}

\newcommand{\highlight}[1]{\textcolor{blue}{#1}}

\newcommand{\proofbox}[1]{\begin{tabular}{l} #1 \end{tabular}}

%%\newcommand{\tarmo}[1]{}
%%\newcommand{\niccolo}[1]{}
%%\newcommand{\cheng}[1]{}

\newcommand{\cheng}[1]{{\color{blue}\textbf{Cheng: }#1}}
\newcommand{\niccolo}[1]{{\color{red}\textbf{Niccol{\`o}: }#1}}
\newcommand{\tarmo}[1]{{\color{ForestGreen}\textbf{Tarmo: }#1}}


\title{Focusing Skew Monoidal Closed Categories}
\author{Tarmo Uustalu
\institute{Reykjavik University, Iceland}
\institute{Tallinn University of Technology, Estonia}
\email{tarmo@ru.is}
\and
Niccol{\`o} Veltri \qquad\qquad Cheng-Syuan Wan
\institute{Tallinn University of Technology, Estonia}
\email{\quad niccolo@cs.ioc.ee \quad\qquad cswan@cs.ioc.ee}
}
\def\titlerunning{Focusing Skew Monoidal Closed Categories}
\def\authorrunning{T. Uustualu, N. Veltri \& C-S. Wan}
\begin{document}
\maketitle
\begin{abstract}
  Monoidal closed categories are used as standard categorical models of intuitionistic noncommutative linear logic: the monoidal unit and tensor model the multiplicative unit and conjunction; the internal hom models linear implication. In recent years, the weaker notion of skew monoidal closed category has been proposed by Ross Street, where the three structural laws of left and right unitality and associativity are not required to be invertible, they are merely natural transformations with a specific orientation. A question arises naturally: Is it possible to find logical proof systems which are naturally modelled by skew monoidal closed categories? We answer positively by introducing three equivalent deductive systems that serve the purpose: a Hilbert-style calculus; a cut-free sequent calculus; a calculus of normal forms obtained from an adaptation of Andreoli's focusing technique to our skew setting. The resulting focused sequent calculus peculiarly employs a system of tags for keeping track of new formulae appearing in the antecedent and appropriately reducing nondeterministic choices in proof search. Focusing solves the coherence problem for skew monoidal closed categories, which implies the existence of effective procedures for deciding equality of canonical morphisms in all models.

%%   In previous works from Uustalu et al, proof theory of some variants of skew monoidal categories are developed.
%%   We also know that given a skew monoidal category, there is a corresponding skew closed structure on the same category.
%%   It motivates us to develop a proof theory of such skew monoidal closed structures and see its meta properties.
%%
%%   In this paper, a Hilbert-style categorical calculus, a cut-free sequent calculus, and a tagged focused sequent calculus are presented.
%%   We prove the soundness and completeness between categorical calculus and sequent calculus.
%%   Moreover any two equivalent proofs from cut-free sequent calculus are mapped into same derivation in tagged focused calculus.
%%   Therefore we have an effective procedure to determine morphism existence and equivalence in skew monoidal closed categories.
\end{abstract}
\section{Introduction}
It is a well-known fact from the late 80s/early 90s that symmetric monoidal closed categories are standard categorical models of MILL, multiplicative intuitionistic linear logic, whose logical connectives comprise of multiplicative unit $\I$ and conjunction $\ot$ and linear implication $\lolli$ \cite{mellies:categorical:09}. Probably lesser-known, though a quite straightforward generalization of the previous sentence which actually predated the inception of linear logic altogether, is the fact that (non-necessarily symmetric) monoidal biclosed categories are categorical models of the noncommutative variant of MILL (NMILL) \cite{abrusci:noncommutative:1990}, where the structural rule of exchange is absent and there are two ordered implications $\lolli$ and $\illol$ \cite{lambek:deductive:68}. A proof system for NMILL is typically called \emph{Lambek calculus}. Lambek refers to the implications $\lolli$ and $\illol$ as residuals, and models of Lambek calculus are thereupon called residual categories. The Lambek calculus is often practically employed in formal investigations of natural languages \cite{lambek:mathematics:58}.
By dropping one of the ordered implication of NMILL one obtains a fragment of the Lambek calculus enjoying categorical semantics in every monoidal closed category.

In recent years, Ross Street introduced the new notion of skew monoidal closed categories \cite{street:skew-closed:2013}. These are a weakening of monoidal closed categories: their structure includes a unit $I$, a tensor $\ot$, an internal hom $\lolli$ and an adjunction relating the latter two operations, as in usual non-skew monoidal categories. The difference lays in the three structural laws of left and right unitality, $\lambda_A : \I \ot A \to A$ and $\rho_A : A \to A \ot \I$, and associativity, $\alpha_{A,B,C} : (A \ot B) \ot C \to A \ot (B \ot C)$, which are usually required to be natural isomorphisms, but in the skew variant are merely natural transformations with the specified orientation. Street originally proposed this weaker notion for better understanding and fixing a famous imbalance, first noticed by Eilenberg and Kelly \cite{eilenberg:closed:1966}, found in the usual adjunction relating monoidal and closed structures \cite{street:skew-closed:2013,uustalu:eilenberg-kelly:2020}. In the last decade, skew monoidal closed categories, together with their non-closed/non-monoidal variants, have been thoroughly studied, with applications ranging from algebra and homotopy theory to programming language semantics \cite{szlachanyi:skew-monoidal:2012,lack:skew:2012,lack:triangulations:2014,altenkirch:monads:2014,buckley:catalan:2015,bourke:skew:2017,bourke:skew:2018,tomita:realizability:21}.

A question arises naturally: is it possible to characterize skew monoidal closed categories as standard categorical models of a certain logical proof system, e.g. a skew variant of NMILL?
This paper provides a positive answer to this question. We introduce a cut-free sequent calculus for a skew variant of NMILL, which we name (rather unoriginally) SkNMILL. Sequents are peculiarly defined as triples $S \mid \Gamma \vdash A$, where the succedent is a single formula $A$ (as in intuitionistic linear logic), but the antecedent is split in two parts: an optional formula $S$, called \emph{stoup} \cite{girard:constructive:91}, and an ordered list of formulae $\Gamma$. Inference rules are similar to the ones of NMILL but with specific structural restrictions for accurately capturing the structural laws of skew monoidal closed categories, and nothing more. In particular, and in analogy with NMILL, the structural rules of contraction, weakening and exchange are all absent. Sets of derivations are quotiented by a congruence relation $\circeq$, carefully chosen to serve as the proof-theoretic counterpart of the equational theory of skew monoidal closed categories. The design of the sequent calculus draws inspiration from, and further advance, the line of work of Uustalu, Veltri and Zeilberger on proof systems for various categories with skew structure: skew semigroup (a.k.a. Tamari order) \cite{zeilberger:semiassociative:19}, skew monoidal (non-closed) \cite{uustalu:sequent:2021,uustalu:proof:nodate} and its symmetric variant \cite{veltri:coherence:2021}, skew prounital (non-monoidal) closed \cite{uustalu:deductive:nodate}.

The metatheory of SkNMILL is developed in two different but related directions:
\begin{enumerate}[($i$)]
  \item We study the categorical semantics of SkNMILL, by showing that the cut-free sequent calculus admits an interpretation of formulae, derivations and equational theory $\circeq$ in every skew monoidal closed category. Moreover, the sequent calculus is \emph{canonical} among skew monoidal closed categories, in the sense that it gives a particular presentation of the \emph{initial} (also called \emph{term} or \emph{free}) model. This can be made precise by introducing a Hilbert-style calculus which directly represents the initial model and subsequently proving that derivations in the two calculi are in a bijective correspondence.
This shows that the categorical semantics of SkNMILL is sound and, most importantly, complete. The latter fact implies that deciding the commutativity of a diagram made of canonical morphisms (i.e. structural laws) in an arbitrary skew monoidal category, a problem typically referred to as \emph{coherence} in the categorical literature, is reduced to checking whether the corresponding derivations in the sequnt calculus are $\circeq$-related.

\item We investigate the proof-thereoretic semantics of SkNMILL, by defining a normalization strategy for sequent calculus derivations wrt. the congruence $\circeq$, when the latter is considered as weakly confluent and strongly normalizing reduction relation. The shape of normal forms is made explicit in a new \emph{focused} sequent calculus, whose derivations act as targets of the normalization procedure. The sequent calculus is ``focused'' in the sense of Andreoli \cite{andreoli:logic:1992}, as it implements a sound and complete goal-directed proof search strategy attempting to build a derivation in the (original, unfocused) sequent calculus. The focused system in this paper extends the previously
developed normal forms for skew monoidal (non-closed) categories \cite{uustalu:sequent:2021,uustalu:proof:nodate}. The presence of both positive ($\I$,$\ot$) and negative ($\lolli$) connectives requires some extra care in the implementation of the proof search algorithm, which is reflected in the design of the focused sequent calculus, in particular when aiming at removing all possible nondeterministic choices (wrt. the conversion $\circeq$) that can arise during proof search. This is technically realized in the focused sequent calculus by the peculiar employment of a system of \emph{tags} for keeping track of new formulae appearing in the antecedent. The focused sequent calculus can also be seen as a concrete presentation of the initial model for SkNMILL, and as such can be used for solving the coherence problem in an effective fashion: deciding commutativity of canonical diagrams in arbitrary skew monoidal categories is equivalent to deciding the \emph{syntactic} equality of the corresponding focused derivations.
\end{enumerate}

%% Srtucture of the paper?
%%
%% Recent discoveries on skew monoidal categories \cite{szlachanyi:skew-monoidal:2012} \cite{lack:skew:2012} \cite{lack:triangulations:2014} (check references in previous papers, at least include papers cited by previous three papers) provide us a good reasons to study their corresponding proof systems.
%% In previous work \cite{uustalu:sequent:2021} \cite{uustalu:deductive:nodate} \cite{uustalu:proof:nodate} \cite{veltri:coherence:2021}, sequent calculus proof system for skew monoidal categories, prounital closed categories, partial normal monoidal categories, and skew symmetric monoidal categories are presented respectively.
%% However, there is still lack of a proof analysis on skew monoidal closed categories.
%% Therefore, in this paper we give sound and complete sequent calculus system with skew monoidal closed categories.
%% Moreover, we use focusing strategy from \cite{andreoli:logic:1992} to solve the coherence problem of skew monoidal closed categories.
%%
%% Interestingly, because of having two connectives in our sequent calculus system, the focusing strategy becomes subtle and involved.
%% It cannot just divide derivations into invertible part and non-invertible part then fix their order.
%% We discovered that it is no harm to arrange invertible rules in a fixed order, but in non-invertible rules, bad thing happens.
%% The naive focused sequent calculus system cannot prove some provable sequents in original sequent calculus.
%% We will see this involved focused sequent calculus system in (?) section.
%%
%% This paper will go through in this order, in section two, we see the Hilbert style calculus of skew monoidal closed categories and its proof equivalences.
%%
%% In third section, a cut-free sequent calculus of skew monoidal closed categories is presented.
%% The key feature is that we have a special formula called stoup at the leftmost position in antecedent of a given sequent.
%% In the same time, for any left rules, it can only be applied to the stoup formula.
%% We will also see preliminarily relationship with Hilbert style calculus.
%%
%% Next section, we will see a focused sequent calculus system and its connection with original sequent calculus system.
%% We will see that all derivations in a same equivalence class under proof conversion relation, they will correspond to a unique derivation in focused system.

\section{A Sequent Calculus for Skew Noncommutative MILL}\label{sec2}
We now introduce the sequent calculus for our skew variant of noncommutative multiplicative linear logic (NMILL), that we call SkNMILL. 

Formulae are inductively generated by the grammar $A,B ::= X \ | \ \I \ | \ A \ot B \ | \ A \lolli B$, where $X$ is an element coming from a fixed set $\mathsf{At}$ of atomic formulae (or propositional variables), $\I$ is a multiplicative unit, $\ot$ is a multiplicative tensor (or conjunction) and $\lolli$ is a linear implication. 

Sequents are triples of the form $S \mid \Gamma \vdash A$, where the succendent $A$ is a single formula (as in NMILL), but the antecedents is divided in two parts: an optional formula $S$, called \emph{stoup} \cite{girard:constructive:91}, and an ordered list of formulae $\Gamma$, called \emph{context}. The peculiar design of sequents, involving the presence of the stoup in the antecedent, comes from previous work on deductive systems with skew structure by Uustalu, Veltri and Zeilberger \cite{uustalu:sequent:2021,uustalu:proof:nodate,uustalu:deductive:nodate,veltri:coherence:2021}.
We reserve the metavariable $S$ to denote a stoup, i.e. $S$ which could be a single formula or empty, in which case we write $S = -$. Atomic formulae will always be names $X,Y,Z$.

Derivations of a sequent $S \mid \Gamma \vdash A$ are inductively generated by the following rules:
%\begin{defn}
%  \emph{SkNMILL} is a sequent calculus system whose sequents are in the form of $S \mid \Gamma \vdash A$ coustructed with following rules
\begin{equation}\label{eq:seqcalc}
  \def\arraystretch{2.5}
  \begin{array}{c}
    \infer[\ax]{A \mid \quad \vdash A}{}
    \quad
    \infer[\pass]{- \mid A , \Gamma \vdash C}{A \mid \Gamma \vdash C}
    \quad
    \infer[\lleft]{A \lolli B \mid \Gamma , \Delta \vdash C}{
      - \mid \Gamma \vdash A
      &
      B \mid \Delta \vdash C
    }
    \quad
    \infer[\lright]{S \mid \Gamma \vdash A \lolli B}{S \mid \Gamma , A \vdash B}
    \\
    \infer[\unitl]{\I \mid \Gamma \vdash C}{- \mid \Gamma \vdash C}
    \quad
    \infer[\tl]{A \ot B \mid \Gamma \vdash C}{A \mid B , \Gamma \vdash C}
    \quad
    \infer[\unitr]{- \mid \quad \vdash \I}{}
    \quad
    \infer[\tr]{S \mid \Gamma , \Delta \vdash A \ot B}{
      S \mid \Gamma \vdash A
      &
      - \mid \Delta \vdash B
    }
  \end{array}
\end{equation}
%\end{defn}


%\niccolo{
%  Explain what the rules do, e.g. when read bottom-up, and what are the differences from the non-skew case
%}

The inference rules in (\ref{eq:seqcalc}) are reminiscent of the ones in the sequent calculus for NMILL (CITE), but there are some crucial differences.
\begin{enumerate}
\item The left logical rules $\unitl$, $\tl$ and $\lleft$, read bottom-up, are only allowed to be applied on the formula in the stoup position. In particular, there is no general way to remove a unit $\I$ nor decompose a tensor $A \ot B$ if these formulae are located in the context and not in the stoup (we will see in (\ref{eq:lleft:gen}) that something can actually be done to deal with implications $A \lolli B$ in the context).
\item The right tensor rule $\tr$, read bottom-up, moves the formula in the stoup, in case such a formula is present, to the stoup of the first premise. In other words, if there is a formula $A'$ in the stoup of the conclusion, after the application of $\tr$ it is not possible to move $A'$ to the antecedent of the second premise.
\item The presence of the stoup implies a distinction between antecedents of the form $A \mid \Gamma$ and $- \mid \Gamma$. The structural rule $\pass$ (for `passivation`), read bottom-up, allows the moving of the leftmost formula in the context to the stoup position, whenever the stoup is initially empty.
\item The logical connectives of NMILL typically include two ordered implications $\lolli$ and $\illol$, which are two variants of linear implication arising from the removal of the exchange rule from intuitionistic linear logic. In SkNMILL only one of the ordered implications is present. It is currently not clear to us whether the inclusion of the second implication to the logic is meaningful and it corresponds to some particular categorical notion.
\end{enumerate}
The restrictions in 1--4 are essential for precisely capturing all the features of skew monoidal closed categories and nothing more, as we discuss in Section \ref{sec:catsem}.
Notice also that, similarly to the case of NMILL, all structural rules of exchange, contraction, and weakening are absent. We give names to derivations and we write $f : S \mid \Gamma \vdash A$ or $S \mid \Gamma \stackrel{f}{\vdash} A$ when $f$ is a particular derivations of the sequent $S \mid \Gamma \vdash A$.

Examples of valid derivations in the sequent calculus, corresponding to structural laws $\lambda$, $\rho$ and $\alpha$ of skew monoidal closed categories (see Definition \ref{def:skewcat}), are given below. %which will help us  understanding the connection with skew monoidal closed categories.
%If we interpret morphism $A \Longrightarrow C$ as a sequent $A \mid \quad \vdash C$, then we can see the natural transformations are derivable in this sequent calculus.
%For example, natural transformations $\lambda : \I \ot A \Longrightarrow A , \rho : A \Longrightarrow A \ot \I$, and $\alpha : (A \ot B) \ot C \Longrightarrow A \ot (B \ot C)$ in skew monoidal closed categories are admissible in SkNMILL:
\begin{equation}\label{eq:lra}
  \begin{array}{c@{\;\quad}cc}
  (\lambda) & (\rho) & (\alpha) \\
  \infer[\tl]{\I \ot A \mid \quad \vdash A}{
    \infer[\unitl]{\I \mid A \vdash A}{
      \infer[\pass]{ \mid A \vdash A}{
        \infer[\ax]{A \mid \quad \vdash A}{}
      }
    }
  }
  &
  \infer[\tr]{A \mid \quad \vdash A \ot \I}{
    \infer[\ax]{A \mid \quad \vdash A}{}
    &
    \infer[\unitr]{- \mid \quad \vdash \I}{}
  }
  &
  \infer[\tl]{(A \ot B) \ot C \mid \quad \vdash A \ot (B \ot C)}{
    \infer[\tl]{A \ot B \mid C \vdash A \ot (B \ot C)}{
      \infer[\tr]{A \mid B , C \vdash A \ot (B \ot C)}{
        \infer[\ax]{A \mid \quad \vdash A}{}
        &
        \infer[\pass]{- \mid B , C \vdash B \ot C}{
          \infer[\tr]{B \mid C \vdash B \ot C}{
            \infer[\ax]{B \mid \quad \vdash B}{}
            &
            \infer[\pass]{- \mid C \vdash C}{
              \infer[\ax]{C \mid \quad \vdash C}{}
            }
          }
        }
      }
    }
  }
  \end{array}
\end{equation}
Examples of non valid sequents include the ``inverses'' of the goals in (\ref{eq:lra}), obtained by swapping the stoup formula with the succedent formula. 
%Another important thing is that SkNMILL could not prove the inverse of any natural transformations above.
The three sequents $X \mid ~ \vdash \I \ot X$, $X \ot \I \mid ~ \vdash X$ and $X \ot (Y \ot Z) \mid ~ \vdash (X \ot Y) \ot Z$ have no derivations. All possible attempts of constructing a valid proof for each of them end in failure.
\begin{displaymath}
  \begin{array}{ccc}
  (\lambda^{-1}) & (\rho)^{-1} & (\alpha^{-1}) \\[6pt]
    \infer[\tr]{X \mid ~\vdash \I \ot X}{
      \deduce[??]{X \mid ~ \vdash \I}{
      }
      &
      \deduce[??]{- \mid ~ \vdash X}{
      }
    }
    &
    \infer[\tl]{X \ot \I \mid \quad \vdash X}{
      \deduce{X \mid \I \vdash X}{??}
    }
    &
    \infer[\tl]{X \ot (Y \ot Z) \mid ~\vdash (X \ot Y) \ot Z}{
      \deduce{X \mid Y \ot Z \vdash (X \ot Y) \ot Z}{??}
    } \\
    (\text{$\tr$ sends $X$ to 1st premise}) &
    (\text{$\unitl$ does not act on $\I$ in context}) &
    (\text{$\tl$ does not act on $\ot$ in context})
  \end{array}
\end{displaymath}
% We interpret $\rho^{-1}$ into $A \ot \I \mid \quad \vdash A$, then according to bottom-up proof search strategy, we first apply $\tl$ but we get stuck immediately.
% Therefore, $\rho^{-1}$ is not derivable in our sequent calculus.
% Other cases are similar.
% In categorical semantics section we will see more details about relationship between SkNMILL and free skew monoidal closed categories.
For analogous reasons, the sequents $\I \lolli A \mid ~ \vdash A$ and $(A \ot B) \lolli C \mid ~ \vdash A \lolli (B \lolli C)$ are valid, while generally their ``inverses'' are not. Moreover, one can turn a derivation of $A \mid ~ \vdash B$ into a derivation of $\I \mid ~ \vdash A \lolli B$, but it is impossible to find a derivation of $X \mid ~ \vdash Y$ from a derivation of $\I \mid ~ \vdash X \lolli Y$.


Sets of derivations are quotiented by a congruence relation $\circeq$, generated by the pairs of derivations in Figure \ref{fig:circeq}. The first three equations are $\eta$-conversions, completely characterizing the $\ax$ rule on non-atomic formulae. The remaining equations are permutative conversions. The congruence $\circeq$ has been carefully chosen to serve as the proof-theoretic counterpart of the equational theory of skew monoidal closed categories, introduced in Definition \ref{def:skewcat}. The subsystem of equations involving only $(\I,\ot)$ originated from \cite{uustalu:sequent:2021}, while the subsystem involving only $\lolli$ comes from \cite{uustalu:deductive:nodate}.
\begin{figure}[t!]
\begin{align*}
  \ax_{\I} &\circeq \unitl \text{ } (\unitr)
  \\
  \ax_{A \ot B} &\circeq \tl \text{ } (\tr \text{ } (\ax_{A} , \pass \text{ } \ax_{B}))
  \\
  \ax_{A \lolli B} &\circeq \lright \text{ } (\lleft \text{ } (\pass \text{ } \ax_{A}, \ax_{B} ))
  \\
  \tr \text{ } (\pass \text{ } f, g) &\circeq \pass \text{ } (\tr \text{ } (f, g)) &&f : A' \mid \Gamma \vdash A, g : - \mid \Delta \vdash B
  \\
  \tr \text{ } (\unitl \text{ } f, g) &\circeq \unitl \text{ } (\tr \text{ } (f , g)) &&f : - \mid \Gamma \vdash A , g : - \mid \Delta \vdash B
  \\
  \tr \text{ } (\tl \text{ } f, g) &\circeq \tl \text{ } (\tr \text{ } (f , g)) &&f : A' \mid B' , \Gamma \vdash A , g : - \mid \Delta \vdash B
  \\
  \pass \text{ } (\lright \text{ } f) &\circeq \lright \text{ } (\pass \text{ } f) &&f : A' \mid \Gamma , A \vdash B
  \\
  \lleft \text{ } (f, \lright \text{ } g) &\circeq \lright \text{ } (\lleft \text{ } (f, g)) &&f : - \mid \Gamma \vdash A', g : B' \mid \Delta , A \vdash B
  \\
  \unitl \text{ } (\lright \text{ } f) &\circeq \lright \text{ } (\unitl \text{ } f) &&f : - \mid \Gamma , A \vdash B
  \\
  \tl \text{ } (\lright \text{ } f) &\circeq \lright \text{ } (\tl \text{ } f) &&f : A \mid B , \Gamma , C \vdash D
  \\
  \lleft \text{ } (f, \tr \text{ } (g, h)) &\circeq \tr \text{ } (\lleft \text{ } (f , g), h) &&f: - \mid \Gamma \vdash A, g : B \mid \Delta \vdash C, h : - \mid \Lambda \vdash D
\end{align*}
\caption{Equivalence of sequent calculus derivations}
\label{fig:circeq}
\end{figure}
%\niccolo{
%We should add all equations, not only the ones about the interaction of monoidal and closed structure. Possibly in a Figure since there are plenty of equations.
%}

The sequent calculus admits cut admissibility.
\begin{theorem}
Two cut rules are admissible %in sequent calculus system for skew monoidal closed categories. $\mathsf{scut}$ and $\mathsf{ccut}$
  \begin{displaymath}
    \infer[\mathsf{scut}]{S \mid \Gamma , \Delta \vdash C}{
      S \mid \Gamma \vdash A
      &
      A \mid \Delta \vdash C
    }
    \quad
    \infer[\mathsf{ccut}]{S \mid \Delta_0 , \Gamma , \Delta_1 \vdash C}{
      - \mid \Gamma \vdash A
      &
      S \mid \Delta_0 , A , \Delta_1 \vdash C
    }
  \end{displaymath}
\end{theorem}
\begin{proof}
  According to previous works \cite{uustalu:sequent:2021} and \cite{uustalu:deductive:nodate} we know two cut rules, $\mathsf{scut}$ and $\mathsf{ccut}$ are admissible in each system separately.
  As our new sequent calculus system is a union of two previous systems, we can just check the $\ot$ and $\lolli$ interaction cases.

  Dealing with $\mathsf{scut}$ first, same as proof strategy in previous papers, we proof by induction on left premise of $\mathsf{scut}$ rule.
  \begin{enumerate}[1. ]
    \item First case is $\mathsf{scut} \text{ } ((\tl \text{ } f), g)$ where $f : A' \mid B' , \Gamma \vdash A, g : A \mid \Delta \vdash C$, then we do subinduction on $g$ to obtain two subcases:
          \begin{enumerate}[a. ]
            \item $g = \lright \text{ } g'$, then we let $\mathsf{scut} \text{ } ((\tl \text{ } f), (\lright \text{ } g')) \defeq \lright \text{ } (\mathsf{scut} \text{ } ((\tl \text{ } f), g'))$.
            \item $g = \lleft \text{ } g'$, then we let $\mathsf{scut} \text{ } ((\tl \text{ } f), (\lright \text{ } g')) \defeq \tl \text{ } (\mathsf{scut} \text{ } (f, (\lright \text{ } g')))$
          \end{enumerate}
    \item $\mathsf{scut} \text{ } ((\lleft f \text{ } (f' , f'')), g)$, where $f' : - \mid \Gamma' \vdash A'$, $f'' : B' \mid \Gamma'' \vdash A$, and $g : A \mid \Delta \vdash C$.
    This case is similar as $\mathsf{scut} \text{ } ((\tl \text{ } f), g)$.
    \item There is no new case for premise is from $\tr$ or $\lright$, because cut formulae are not matched.
  \end{enumerate}
  Next we see the cut-freeness of $\mathsf{ccut}$ rule.
  Similarly, we prove it by induction on the second premise.
  \begin{enumerate}[1. ]
    \item $\mathsf{ccut} \text{ } (f , (\lleft \text{ } (g, h)))$ where $f : - \mid \Gamma \vdash A$, $g : - \mid \Delta_1 \vdash A'$, $h : B' \mid \Delta_2 \vdash C$, then we do subinduciton on $f$:
    \begin{enumerate}[a. ]
      \item The only possibility is $A = A' \ot B'$ and $f = \tr (f_1, f_2)$ where $f_1 : - \mid \Gamma_1 \vdash A'$, $f_2 : - \mid \Gamma_2 \vdash B'$, then depending on $A' \ot B'$ in $\Delta_1$ or $\Delta_2$,
      we let $\mathsf{ccut} \text{ } (f , (\lleft \text{ } (g, h))) \defeq \lleft \text{ } (\mathsf{ccut} \text{ } (f, g), h)$ or $ \lleft \text{ } (g, (\mathsf{ccut} \text{ } (f, h)))$, respectively.
    \end{enumerate}
    \item $\mathsf{ccut} \text{ } (f , (\tr \text{ } (g, h)))$ where $f : - \mid \Gamma \vdash A$, $g : S \mid \Delta_1 \vdash A'$, $h : - \mid \Delta_2 \vdash B'$ is similar as above.
    \item $\mathsf{ccut} \text{ } (f, (\tl \text{ } g))$ and
    $\mathsf{ccut} \text{ } (f, (\lright \text{ } h))$ cases are similar, we permute $\mathsf{ccut}$ up,
    where $g : A' \mid B' , \Delta_0 , A , \Delta_1 \vdash C$, $h : S \mid \Delta_0 , A , \Delta_1 , A' \vdash B'$.
  \end{enumerate}
\end{proof}
The two cut rule satisfy a large number of
$\circeq$-equations, see e.g.~\cite[Figures 5 and 6]{uustalu:sequent:2021} for the
list of such equations not involving $\lolli$.
Other important admissible rules are listed below.
\begin{itemize}
\item The left rules for $\I$ and $\ot$ are invertible wrt. $\circeq$, and similarly the right rule for $\lolli$. This in particular implies the admissibility of the following rules:
\begin{displaymath}
  \infer[\tl^{-1}]{A \mid B, \Gamma \vdash C}{A \ot B \mid \Gamma \vdash C}
  \qquad
  \infer[\unitl^{-1}]{- \mid \Gamma \vdash C}{\I \mid \Gamma \vdash C}
  \qquad
  \infer[\lright^{-1}]{S \mid \Gamma , A \vdash B}{S \mid \Gamma \vdash A \lolli B}
\end{displaymath}
Notice that no other rule is invertible, and in particular not the
passivation rule $\pass$.

\item Applications of the invertible left logical rules can be iterated, and similarly for the invertible right $\lright$ rule, resulting in the two admissible rules
\begin{equation}\label{eq:inter:ante}
  \infer[\mathsf{L}^\star]{\ldbc S \mid \Gamma \rdbc_{\ot} \mid \Delta \vdash C}{
    S \mid \Gamma ,\Delta \vdash C
  }
  \qquad
  \infer[\lright^\star]{S \mid \Gamma \vdash \ldbc \Delta \mid C \rdbc_{\lolli}}{
    S \mid \Gamma , \Delta \vdash C
  }
\end{equation}
The interpretation of antecedents $\ldbc S \mid \Gamma \rdbc_{\ot}$ in (\ref{eq:inter:ante}) is the formula obtained by subtituting the separator $\mid$ and the commas with tensors: $\ldbc S \mid A_1,\dots,A_n \rdbc_{\ot} = (\dots ((\ldbc S \llangle \ot A_1) \ot A_2) \dots ) \ot A_n$, where $\ldbc - \llangle = \I$ and $\ldbc A \llangle = A$. 
% as $\ldbc S \mid \Gamma \rdbc_{\ot} = \ldbc S \llangle \text{ } \llangle \Gamma \rdbc_{\ot}$, where
% \begin{align*}
%   \ldbc - \llangle = \I \qquad \ldbc A \llangle = A \quad\qquad A \llangle \text{ } \rdbc_{\ot} = A \qquad A \llangle B , \Gamma \rdbc_{\ot} = (A \ot B ) \llangle \Gamma \rdbc_{\ot}
% \end{align*}
% In other words, $\ldbc S \mid A_1,\dots,A_n \rdbc_{\ot} = (\dots ((\ldbc S \llangle \ot A_1) \ot A_2) \dots ) \ot A_n$. 
Dually, the formula $\ldbc \Delta \mid C \rdbc_{\lolli}$ in (\ref{eq:inter:ante}) is obtained by substituting $\mid$ and commas with implications: 
% \begin{displaymath}
%   \ldbc \quad \mid C \rdbc_{\lolli} = C \qquad \ldbc A , \Gamma \mid C \rdbc_{\lolli} = A \lolli \ldbc \Gamma \mid C \rdbc_{\lolli}
% \end{displaymath}
$\ldbc A_1,\dots,A_n \mid C \rdbc_{\lolli} = A_1 \lolli (A_2 \lolli (\dots \lolli (A_n \lolli C)))$.

\item A second left implication rule, acting on a formula $A \lolli B$ in the context, is derivable using cut:
\begin{equation}\label{eq:lleft:gen}
    \!\!\!
  \proofbox{
    \infer[\lleft_{\mathsf{C}}]{S \mid \Delta_0, A \lolli B, \Gamma , \Delta_1 \vdash C}{
      - \mid \Gamma \stackrel{f}{\vdash} A
      &
      S \mid \Delta_0, B , \Delta_1 \stackrel{g}{\vdash} C
    }
  }
  {=}
  \proofbox{
    \infer[\mathsf{ccut}]{S \mid \Delta_0, A \lolli B, \Gamma , \Delta_1 \vdash C}{
      \infer[\pass]{- \mid A \lolli B, \Gamma \vdash B}{
        \infer[\lleft]{A \lolli B \mid \Gamma \vdash B}{
          - \mid \Gamma \stackrel{f}{\vdash} A
          &
          \infer[\ax]{B \mid ~ \vdash B}{}
        }
      }
      &
      S \mid \Delta_0, B , \Delta_1 \stackrel{g}{\vdash} C
    }
  }
\end{equation}
\end{itemize}

% There are some useful rules are admissible in SkNMILL.
% \begin{remark}
%   Following rules are admissible in SkNMILL:
%   \begin{displaymath}
%     \infer{\ldbc S \mid \Gamma \rdbc_{\ot} \mid \Delta \vdash C}{
%       S \mid \Gamma ,\Delta \vdash C
%     }
%     \qquad
%     \infer{S \mid \Gamma \vdash \ldbc \Delta \mid C \rdbc_{\lolli}}{
%       S \mid \Gamma , \Delta \vdash C
%     }
%     \qquad
%     \infer{S \mid \Delta_0, A \lolli B, \Gamma , \Delta_1 \vdash C}{
%       - \mid \Gamma \vdash A
%       &
%       S \mid \Delta_0, B , \Delta_1 \vdash C
%     }
%   \end{displaymath}
%   Inductive definitions of formulae $\ldbc A \mid \Gamma \rdbc_{\ot}$ (defined as $\ldbc S \mid \Gamma \rdbc_{\ot} = \ldbc S \llangle \text{ } \llangle \Gamma \rdbc_{\ot}$) and $\ldbc \Gamma \mid C \rdbc_{\lolli}$ are originated from \cite{uustalu:sequent:2021} and \cite{uustalu:deductive:nodate} respectively:
%   \begin{align*}
%     \ldbc - \llangle = \I \qquad \ldbc A \llangle = A \quad\qquad A \llangle \text{ } \rdbc_{\ot} = A \qquad A \llangle B , \Gamma \rdbc_{\ot} = (A \ot B ) \llangle \text{ } \Gamma \rdbc_{\ot}
%   \end{align*}
%   \begin{displaymath}
%     \ldbc \quad \mid C \rdbc_{\lolli} = C \qquad \ldbc A , \Gamma \mid C \rdbc_{\lolli} = A \lolli \ldbc \Gamma \mid C \rdbc_{\lolli}
%   \end{displaymath}
% \end{remark}

%SkNMILL is a cut-free calculus so we are in a good shape and could discuss more about its correspondence with categorical structures.

%\niccolo{The above proof is interesting, but also quite long. We could move this to the Appendix if we need to make some room in the main text.
%}

%\niccolo{
%  We should mention (in a Proposition?) that some other rules are admissible: inverses of $\unitl$, $\tl$ and $\lright$ (and these are the only invertible ones, in particular $\pass$ is not invertible), and also:
%  \[
%  \infer{\ldbc S \mid \Gamma \rdbc_{\ot} \mid \Delta \vdash C}{
%    S \mid \Gamma ,\Delta \vdash C
%  }
%  \qquad
%  \infer{S \mid \Gamma \vdash \ldbc \Delta \mid C \rdbc_{\lolli}}{
%    S \mid \Gamma , \Delta \vdash C
%  }
%  \qquad
%    - \mid \Gamma \vdash A
%    &
%    S \mid \Delta_0, B , \Delta_1 \vdash C
%  }
%  \]
%  The last one using cut. So we could introduce the definitions of $\ldbc S \mid \Gamma \rdbc_{\ot}$ and $\ldbc \Delta \mid C \rdbc_{\lolli}$ at this point.
%  }

%%We have a cut-free proof theory which characterizes skew monoidal closed categories nicely so far.
%%According to Curry-Howard-Lambek correspondence, we can interpret morphisms in skew monoidal closed categories as derivations in the sequent calculus system.
%%This fact allows us to tackle problems in skew monoidal closed categories by methods in proof theory.
%%In particular, we will use Andreoli's focusing technique \cite{andreoli:logic:1992} to analyze and solve coherence problem of skew monoidal closed categories.

\section{Categorical Semantics via Skew Monoidal Closed Categories}
\label{sec:catsem}

%\niccolo{
%I don't think we should give the definition of skew monoidal closed category using all these structural laws and equations, since we know that it can be done with much less (and all this structure takes a lot of room!). I suggest following the definition given in \cite{uustalu:eilenberg-kelly:2020}, using $\lambda$, $\rho$ and $\alpha$ (it is enough to state that there is an adjunction, no need to show $\eta/\epsilon$ nor $p/\mathsf{M}$). Afterwards we can comment on the fact that the definition can equivalently be stated using $i$, $j$ and $L$, and that the adjunction cannot be generally internalized.
%}

%\niccolo{
%I suggest following the discussion in the introduction and proceed as follows:
%\begin{enumerate}
%\item The sequent calculus defines a skew monoidal closed category FSkMCC($\mathsf{At}$) with formulae as objects and set of derivations $A \mid ~ \vdash B$ quotiented by $\circeq$ as homset between objects $A$ and $B$. We already mentioned in the previous section the definitions of $\lambda$, $\rho$ and $\alpha$, I am not sure if that discussion should be moved here.
%\item Given a skew monoidal closed category $\mathbb{C}$, every function $F_{\mathsf{At}} : \mathsf{At} \to \mathbb{C}$, giving an interpretation of atomic formulae as objects of $\mathbb{C}$, extends to a skew monoidal closed functor $F :$ FSkMCC$(\mathsf{At}) \to \mathbb{C}$ (then we should give the definition of skew monoidal closed functor beforehand). This shows that skew monoidal closed categories are categorical models of SkNMILL.
%\item FSkMCC($\mathsf{At}$) is the initial model, a.k.a. the free skew monoidal closed category on $\mathsf{At}$. The proof of this calculus is where we should introduce the Hilbert-style calculus (again with a reduced number of rules/equations). Theorem \ref{thm:wkcomplete} is a consequence of point 1 above, i.e. FSkMCC($\mathsf{At}$) is a skew monoidal closed category. Theorem  \ref{thm:sound} instead follows from point 2.
%\end{enumerate}
%}
\begin{defn}\label{def:skewcat}
A \emph{skew monoidal closed category} $\mathbb{C}$ is a category with a unit object $\I$ and two functors $\ot : \mathbb{C} \times \mathbb{C} \rightarrow \mathbb{C}$ and $\lolli : \mathbb{C}^{\mathsf{op}} \times \mathbb{C} \rightarrow \mathbb{C}$ forming an adjunction $- \ot B \dashv B \lolli -$ natural in $B$, 
and three natural transfomations %$\lambda$, $\rho$, $\alpha$ with components
\begin{displaymath}
  \lambda_A : \I \ot A \to A \qquad \rho_A : A \to A \ot \I \qquad \alpha_{A,B,C} : (A \ot B) \ot C \to A \ot (B \ot C)
\end{displaymath}
satisfying the following axioms due to Mac Lane \cite{maclane1963natural}:
\begin{center}
  %(m1)
  % https://q.uiver.app/?q=WzAsMyxbMSwwLCJcXEkgXFxvdCBcXEkiXSxbMCwxLCJcXEkiXSxbMiwxLCJcXEkiXSxbMSwwLCJcXHJob197XFxJfSJdLFswLDIsIlxcbGFtYmRhX3tcXEl9Il0sWzEsMiwiIiwyLHsibGV2ZWwiOjIsInN0eWxlIjp7ImhlYWQiOnsibmFtZSI6Im5vbmUifX19XV0=
\begin{tikzcd}
	& {\I \ot \I} \\
	\I && \I
	\arrow["{\rho_{\I}}", from=2-1, to=1-2]
	\arrow["{\lambda_{\I}}", from=1-2, to=2-3]
	\arrow[Rightarrow, no head, from=2-1, to=2-3]
\end{tikzcd}
%(m2)
% https://q.uiver.app/?q=WzAsNCxbMCwwLCIoQSBcXG90IFxcSSkgXFxvdCBCIl0sWzEsMCwiQSBcXG90IChcXEkgXFxvdCBCKSJdLFsxLDEsIkEgXFxvdCBCIl0sWzAsMSwiQSBcXG90IEIiXSxbMywyLCIiLDAseyJsZXZlbCI6Miwic3R5bGUiOnsiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dLFszLDAsIlxccmhvX0EgXFxvdCBCIl0sWzEsMiwiQSBcXG90IFxcbGFtYmRhX3tCfSJdLFswLDEsIlxcYWxwaGFfe0EgLCBcXEkgLCBCfSJdXQ==
\begin{tikzcd}
	{(A \ot \I) \ot B} & {A \ot (\I \ot B)} \\
	{A \ot B} & {A \ot B}
	\arrow[Rightarrow, no head, from=2-1, to=2-2]
	\arrow["{\rho_A \ot B}", from=2-1, to=1-1]
	\arrow["{A \ot \lambda_{B}}", from=1-2, to=2-2]
	\arrow["{\alpha_{A , \I , B}}", from=1-1, to=1-2]
\end{tikzcd}

%(m3)
% https://q.uiver.app/?q=WzAsMyxbMCwwLCIoXFxJIFxcb3QgQSApIFxcb3QgQiJdLFsyLDAsIlxcSSBcXG90IChBIFxcb3QgQikiXSxbMSwxLCJBIFxcb3QgQiJdLFswLDEsIlxcYWxwaGFfe1xcSSAsIEEgLEJ9Il0sWzEsMiwiXFxsYW1iZGFfe0EgXFxvdCBCfSJdLFswLDIsIlxcbGFtYmRhX3tBfSBcXG90IEIiLDJdXQ==
\begin{tikzcd}
	{(\I \ot A ) \ot B} && {\I \ot (A \ot B)} \\
	& {A \ot B}
	\arrow["{\alpha_{\I , A ,B}}", from=1-1, to=1-3]
	\arrow["{\lambda_{A \ot B}}", from=1-3, to=2-2]
	\arrow["{\lambda_{A} \ot B}"', from=1-1, to=2-2]
\end{tikzcd}
%(m4)
% https://q.uiver.app/?q=WzAsMyxbMCwwLCIoQSBcXG90IEIpIFxcb3QgXFxJIl0sWzIsMCwiQSBcXG90IChCIFxcb3QgXFxJKSJdLFsxLDEsIkEgXFxvdCBCIl0sWzAsMSwiXFxhbHBoYV97QSAsIEIsIFxcSX0iXSxbMiwxLCJBIFxcb3QgXFxyaG9fQiIsMl0sWzIsMCwiXFxyaG9fe0EgXFxvdCBCfSJdXQ==
\begin{tikzcd}
	{(A \ot B) \ot \I} && {A \ot (B \ot \I)} \\
	& {A \ot B}
	\arrow["{\alpha_{A , B, \I}}", from=1-1, to=1-3]
	\arrow["{A \ot \rho_B}"', from=2-2, to=1-3]
	\arrow["{\rho_{A \ot B}}", from=2-2, to=1-1]
\end{tikzcd}

%(m5)
% https://q.uiver.app/?q=WzAsNSxbMCwwLCIoQVxcb3QgKEIgXFxvdCBDKSkgXFxvdCBEIl0sWzIsMCwiQSBcXG90ICgoQiBcXG90IEMpIFxcb3QgRCkiXSxbMiwxLCJBIFxcb3QgKEIgXFxvdCAoQyBcXG90IEQpKSJdLFsxLDEsIihBIFxcb3QgQikgXFxvdCAoQyBcXG90IEQpIl0sWzAsMSwiKChBIFxcb3QgKEJcXG90IEMpIFxcb3QgRCkiXSxbMCwxLCJcXGFscGhhX3tBICwgQlxcb3QgQyAsIER9Il0sWzEsMiwiQSBcXG90IFxcYWxwaGFfe0IgLCBDICxEfSJdLFszLDIsIlxcYWxwaGFfe0EgLEIgLENcXG90IER9IiwyXSxbNCwzLCJcXGFscGhhX3tBIFxcb3QgQiAsIEMgLCBEfSIsMl0sWzQsMCwiXFxhbHBoYV97QSAsIEIgLEN9IFxcb3QgRCJdXQ==
\begin{tikzcd}
	{(A\ot (B \ot C)) \ot D} && {A \ot ((B \ot C) \ot D)} \\
	{((A \ot (B\ot C) \ot D)} & {(A \ot B) \ot (C \ot D)} & {A \ot (B \ot (C \ot D))}
	\arrow["{\alpha_{A , B\ot C , D}}", from=1-1, to=1-3]
	\arrow["{A \ot \alpha_{B , C ,D}}", from=1-3, to=2-3]
	\arrow["{\alpha_{A ,B ,C\ot D}}"', from=2-2, to=2-3]
	\arrow["{\alpha_{A \ot B , C , D}}"', from=2-1, to=2-2]
	\arrow["{\alpha_{A , B ,C} \ot D}", from=2-1, to=1-1]
\end{tikzcd}
\end{center}
\end{defn}
The notion of skew monoidal closed category admits other equivalent characterizations \cite{street:skew-closed:2013,uustalu:eilenberg-kelly:2020}. Tuples of natural transformations $(\lambda , \rho , \alpha)$ are in bijective correspondence with tuples 
\[
j_A : \I \to A \lolli A \qquad i_A : \I \lolli A \to A \qquad L_{A,B,C} : B \lolli C \to (A \lolli B) \lolli (A \lolli C)
\]
Moreover, $\alpha$ and $L$ are interdefinable with a natural transformation $\mathsf{p}_{A , B , C} : (A \ot B) \lolli C \Longrightarrow A \lolli (B \lolli C)$, representing an internalization of the adjunction between $\ot$ and $\lolli$.


\begin{defn}
  A \emph{(strict) skew monoidal closed functor} $F : \mathbb{C} \rightarrow \mathbb{D}$ between skew monoidal closed categories $(\mathbb{C} , \I , \ot , \lolli)$ and $(\mathbb{D} , \I' , \ot' , \lolli')$ is a functor from $\mathbb{C}$ to $\mathbb{D}$ satisfying
%   \begin{align*}
%     F \I \cong \I' \qquad F (C_{1} \ot C_{2}) \cong F C_{1} \ot' F C_{2} \qquad
%     F(C_{1} \lolli C_{2}) \cong F C_{1} \lolli' F C_{2}
%   \end{align*}
%   and preserving strutural laws $\lambda$, $\rho$, and $\alpha$ and .
%   \\
%   A skew monoidal closed functor is strict if
%  \begin{align*}
    $F \I = \I'$, $F (A \ot B) = F A \ot' F B$ and 
    $F(A \lolli B) = F A \lolli' F B$, also preserving  strutural laws $\lambda$, $\rho$, and $\alpha$ on the nose.
%  \end{align*}
\end{defn}
%Next we show that given any skew monoidal closed category $\mathbb{C}$ and a function $F_{\mathsf{At}} : \mathsf{At} \rightarrow \mathbb{C}$ providing interpretation of atomic formulae as objects of $\mathbb{C}$, we can extend function $F_{\mathsf{At}}$ to a skew monoidal closed functor $F : \text{FSkMCC} \rightarrow \mathbb{C}$.

\begin{example}
A good example??
\end{example}

The formulae, derivations and equivalence $\circeq$ of the sequent calculus for SkNMILL determine a skew monoidal closed category FSkMCC($\mathsf{At}$). Objects are formulae, while the set of morphisms between formulaa $A$ and $B$ is the set of derivations $A \mid ~ \vdash B$ quotiented by the equivalence relation $\circeq$. The identity morphism is $\ax$, while $\mathsf{scut}$ gives composition of morphisms. The construction of derivations for the structural laws $\lambda$, $\rho$,$\alpha$ is given in (\ref{eq:lra}). It is also possible to prove that the relation $\sim$ satisfies the 5 Mac Lane axioms.


Given any skew monoidal closed category $(\mathbb{C} , \I' , \ot' , \lolli')$ and a function $F_{\mathsf{At}} : \mathsf{At} \rightarrow \mathbb{C}$ providing interpretation of atomic formulae as objects of $\mathbb{C}$, there exists a skew monoidal closed functor $F : \text{FSkMCC}(\mathsf{At}) \rightarrow \mathbb{C}$.

  Let $\mathsf{Fma}$ denote the collection of FSkMCC($\mathsf{At}$) and $\mathbb{C}_{0}$ denote the collection of $\mathbb{C}$, then we can define a mapping $F_{0} : \mathsf{Fma} \rightarrow \mathbb{C}_{0}$
  \begin{equation*}
    F_{0} (A)=
    \begin{cases}
      F_{\mathsf{At}} (A) &\text{if } A \text{ is atomic}
      \\
      \I' &\text{if } A = \I
      \\
      F_{\mathsf{At}} (B) \ot' F_{\mathsf{At}} (C) &\text{if } A = B \ot C
      \\
      F_{\mathsf{At}} (B) \lolli' F_{\mathsf{At}} (C) &\text{if } A = B \lolli C
    \end{cases}
  \end{equation*}
  and another mapping $F_{1} : A \mid \quad \vdash B \to \mathbb{C} (F_{0} (A), F_{0} (B))$ inductively.
  Integrating mapping $F_{0}$ and $F_{1}$, we have a skew monoidal closed functor from FSkMCC($\mathsf{At}$) to $\mathbb{C}$.
  This fact tells us that any skew monoidal closed category is also a categorical model of SkMNILL.

  Next we consider the free construction of skew monoidal closed category over a set $\mathsf{At}$ with Hilbert-style presentation.
  \begin{defn}
    A free skew monoidal closed category FSkMCC($\mathsf{At}$)$_{H}$ whose objects is $\mathsf{Fma}$ and derivations are sequents $A \Longrightarrow C$ is a Hilbert-style sequent calculus system (categorical calculus) constructed from following inference rules
    \begin{displaymath}
        \infer[\id]{A \Longrightarrow A}{}
        \quad
        \infer[\mathsf{comp}]{A \Longrightarrow C}{
          A \Longrightarrow B
          &
          B \Longrightarrow C
        }
    \end{displaymath}
    \begin{displaymath}
      \infer[\otimes]{A \ot B \Longrightarrow C \ot D}{
        A \Longrightarrow C
        &
        B \Longrightarrow D
      }
      \quad
      \infer[\lolli]{A \lolli B \Longrightarrow C \lolli D}{
        C \Longrightarrow A
        &
        B \Longrightarrow D
      }
    \end{displaymath}
    \begin{displaymath}
      \infer[\lambda]{\I \ot A \Longrightarrow A}{}
      \quad
      \infer[\rho]{A \Longrightarrow A \ot \I}{}
      \quad
      \infer[\alpha]{(A \ot B) \ot C \Longrightarrow A \ot (B \ot C)}{}
    \end{displaymath}
    \begin{displaymath}
      \infer[\epsilon]{(A \lolli B) \ot A \Longrightarrow B}{}
      \quad
      \infer[\mathsf{adj_1}]{A \Longrightarrow B \lolli C}{A \ot B \Longrightarrow C}
    \end{displaymath}
    satisfying following equations
    \begin{align*}
      \id \circ f \doteq f \qquad f \doteq f \circ \id \qquad (f \circ g) \circ h \doteq f \circ (g \circ h) &&\text{(category laws)}
      \\
      \id \ot \id \doteq \id \qquad (h \circ f) \ot (k \circ g) \doteq h \ot k \circ f \ot g &&(\ot \text{ functorial})
      \\
      \id \lolli \id \doteq \id \qquad (h \circ f) \ot (k \circ g) \doteq f \lolli k \circ h \lolli g &&(\lolli \text{ functorial})
    \end{align*}
    \begin{align*}
        \lambda \circ \id \ot f \doteq f \circ \lambda
        \\
        \rho \circ f \doteq f \ot \id \circ \rho &&\lambda , \rho , \alpha \text{ are natural transformations}
        \\
        \alpha \circ (f \ot g) \ot h \doteq f \ot (g \ot h) \circ \alpha
        %%%%Extra natural version?
        %f \lolli (g \lolli h) \circ p \doteq p \circ (f \ot g) \lolli h
        %\\
        %f \lolli h \circ \mathsf{M} \doteq \mathsf{M} \circ (g \lolli h) \ot (f \lolli g)
        %%%%Extra natural version?
    \end{align*}
    \begin{align*}
        \lambda \circ \rho \doteq \id \qquad \id \doteq \id \ot \lambda \circ \alpha \circ \rho \ot \id
        \\
        \lambda \circ \alpha \doteq \lambda \ot \id \qquad \alpha \circ \rho \doteq \id \ot \rho
        &&\text{(m1)-(m5)}
        \\
        \alpha \circ \alpha \doteq
        \id \ot \alpha \circ \alpha \circ \alpha \ot \id
    \end{align*}
  \end{defn}
\begin{fact}
  Natural transformations $i$, $j$, $\msfL$, $\mathsf{p}$, and $\eta : A \Longrightarrow B \lolli (A \ot B)$ and inference rule
  \begin{displaymath}
    \infer[\mathsf{adj}_{2}]{A \ot B \Longrightarrow C}{A \Longrightarrow B \lolli C}
  \end{displaymath}
   are interdefinable with $\lambda$, $\rho$, $\alpha$, $\epsilon$, and $\mathsf{adj}_{1}$.
  \begin{align*}
    i &\defeq \mathsf{comp} \text{ } (\rho ,  (\mathsf{adj_2} \text{ } \id))
    \\
    j &\defeq \mathsf{adj_1} \text{ } \lambda
    \\
    \msfL &\defeq \mathsf{adj_1} \text{ } (\mathsf{adj_1} \text{ } (\mathsf{comp} \text{ } (\alpha , (\mathsf{comp} \text{ } (\id \ot \epsilon) , \epsilon))))
    \\
    \eta &\defeq \mathsf{adj_1} \text{ } \id
    \\
    \mathsf{adj_2} &\defeq \comp \text{ } ((f \ot \id), \epsilon) && f : A \Longrightarrow B \lolli C
    \\
    \mathsf{p} &\defeq \comp \text{ } (\msfL , (\eta \lolli \id))
    %\\
    %\mathsf{M} &\defeq \comp \text{ } ((\msfL \ot \id) , \epsilon)
  \end{align*}
\end{fact}
Soundness and completeness among FSkMCC($\mathsf{At}$)$_{H}$ and FSkMCC($\mathsf{At}$) is justified because we can prove bijection of derivations (quotiented by $\doteq$ and $\circeq$ respectively) between these two deductive systems by induction on derivations.
\section{Proof-Theoretic Semantics via Focusing}
\begin{defn}
  The \emph{focused sequent calculus} for SkNMILL has four phases($\RI , \LI , \Pass$, and $\F$) with  special annotations $\bullet$(tag) on $\vdash$ so SkNMILL defines a skew monoidal closed category FSkMCC($\mathsf{At}$).

  \begin{displaymath}
    \infer[\lright]{S \mid \Gamma \vdash^{x}_{\RI} A \lolli B}{S \mid \Gamma , A^{x} \vdash^{x}_{\RI} B}
    \quad
    \infer[\LI 2 \RI]{S \mid \Gamma \vdash^{x}_{\RI} P}{S \mid \Gamma \vdash^{x}_{\LI} P}
  \end{displaymath}
  \begin{displaymath}
    \infer[\unitl]{\I \mid \Gamma \xvdash_{\LI} P}{- \mid \Gamma \xvdash_{\LI} P}
    \quad
    \infer[\tl]{A \ot B \mid \Gamma \xvdash_{\LI} P}{A \mid B , \Gamma \xvdash_{\LI} P}
    \quad
    \infer[\Pass 2 \LI]{T \mid \Gamma \xvdash_{\LI} P}{T \mid \Gamma \xvdash_{\Pass} P}
  \end{displaymath}
  \begin{displaymath}
    \infer[\pass]{- \mid A^{x} , \Gamma \xvdash_{\Pass} P }{A^{\circ} \mid \Gamma \xvdash_{\LI} P}
    \quad
    \infer[\F 2 \Pass]{T \mid \Gamma \xvdash_{\Pass} P}{T \mid \Gamma \xvdash_{\F} P}
  \end{displaymath}
  \begin{displaymath}
    \infer[\tr]{T \mid \Gamma , \Delta \xvdash_{\F} A \ot B}{
      T \mid \Gamma \vdash^{\bullet}_{\RI} A
      &
      - \mid \Delta \vdash_{\RI} B
    }
    \quad
    \infer[\lleft]{A \lolli B \mid \Gamma , \Delta \xvdash_{\F} P}{
      - \mid \Gamma^{\circ} \vdash_{\RI} A
      &
      B \mid \Delta \vdash_{\LI} P
      &
      \text{if } x = \bullet \text{, then } \exists F^{\bullet} \in \Gamma
    }
  \end{displaymath}
  \begin{displaymath}
    \infer[\ax]{A \mid \quad \vdash_{\F} A}{}
    \quad
    \infer[\unitr]{- \mid \quad \vdash \I}{}
  \end{displaymath}
\end{defn}

We explain the intuition of this focused system from bottom-up proof search perspective.
One important thing is that even though we have tagged suquents in a proof, but the end sequent should keep tag free.
\begin{itemize}
  \item In phase $\RI$ (`for right invertible'), only right invertible rules can be applied, in particular, $\lright$.
  If conclusion of a $\RI$ application is black, then we add a tag to keep tracking the formulae decomposed from $\RI$.
  In the end, we decompose $C$ until it becomes a positive $P$ where $P \neq A \lolli B$, then move to phase $\LI$.
  \item Next, in phase $\LI$ (`for left invertible') we destruct $S$ by left invertible rules $\tl$ and $\unitl$ until we obtain an irreducible stoup $S'$ where $S'$ is not in $A' \ot B'$ or $\I$.
  In this phase we do not need to enforce the order between $\unitl$ and $\tl$ applicaitons because they are not exchangeable.
  \item Then we move to passivation phase $\Pass$, where only $\pass$ can be applied to sequents.
  Notice that for the active formula in each application of $\pass$, we remove its tag and the sequent goes back to phase $\LI$ to decompose stoup formula again because the passviated formula $A$ could be reducible.
  \item In phase $\F$ (for `focusing'), we have four rules, $\tr$, $\lleft$, $\ax$, and $\unitr$.
  Here we have special resrictions on $\tl$ and $\lleft$.
  In general, we prefer to decompose $\lolli$ in stoup prior then $\ot$ in conclusion.
  However, if there are some formula are packed into conclusion, then whole bottom-up proof strategy fails.
  We will see comparison between involved and naive focused system later.
  For rule $\tr$, we let its left and right premises back to phase $\RI$ because $A$ and $B$ could be negative formulas.
  Especially, the left premise becomes a tagged sequent.
  In a $\lleft$ application, if conclusion sequent is tagged, then at least exists a formula $F$ is black in the context of left premise.
  Remembering that we prefer $\lleft$ prior to $\tr$, so if there is any $\tr$ before $\lleft$, then we have to ensure that we cannot do $\lleft$ first.
  Black tags produced by $\lright$ play essential roles here because if there is no black formula in $\Gamma$, then it implies that we cannot do $\lleft$ after $\tr$.
  In this situation, we must decompose $\lolli$ in stoup prior than $\ot$ in conclusion.
\end{itemize}

Assume that we would prefer to decompose $\lolli$ in stoup prior than $\ot$ in succedent without tag setting above.
In this naive focused sequent calculus, we cannot prove $X \lolli Y \mid Z \vdash_{\RI} (X \lolli Y) \ot Z$, where $X, Y$, and $Z$ are atomic formulae:
\begin{displaymath}
  \infer[\mathsf{\LI 2 \RI}]{X \multimap Y \mid Z \vdash_{\mathsf{RI}} (X\multimap Y) \otimes Z}{
    \infer[\Pass 2 \LI]{X \multimap Y \mid Z \vdash_{\mathsf{LI}} (X\multimap Y) \otimes Z}{
      \infer[\F 2 \Pass]{X \multimap Y \mid Z \vdash_{\Pass} (X\multimap Y) \otimes Z}{
        \infer[\lleft]{X \multimap Y \mid Z \vdash_{\F} (X\multimap Y) \otimes Z}{
          \deduce{- \mid \quad \vdash_{\RI} X}{??}
          &
          \infer[\Pass 2 \LI]{Y \mid Z \vdash_{\LI} (X \lolli Y) \ot Z}{
            \infer[\F 2 \Pass]{Y \mid Z \vdash_{\Pass} (X \lolli Y) \ot Z}{
              \infer[\tr]{Y \mid Z \vdash_{\F} (X \lolli Y) \ot Z}{
                \infer[\lright]{Y \mid \quad \vdash_{\RI} X \lolli Y}{
                  \deduce{Y \mid X \vdash_{\RI} Y}{??}
                }
                &
                \infer[\LI 2 \RI]{- \mid Z \vdash_{\RI} Z}{
                  \infer[\Pass 2 \LI]{- \mid Z \vdash_{\LI} Z}{
                    \infer[\pass]{- \mid Z \vdash_{\Pass} Z}{
                      \infer[\Pass 2 \LI]{Z \mid \quad \vdash_{\LI} Z}{
                        \infer[\F 2 \Pass]{Z \mid \quad \vdash_{\Pass} Z}{
                          \infer[\ax]{Z \mid \quad \vdash_{\F} Z}{}
                        }
                      }
                    }
                  }
                }
              }
            }
          }
        }
      }
    }
  }
\end{displaymath}
The key point is that $X$ in $(X \lolli Y) \ot Z$ is locked, so that we could not move it to the context of left premise to close proof tree.
However, in tagged setting, we can avoid this awkward situation:
\begin{displaymath}
  \infer[\mathsf{\LI 2 \RI}]{X \multimap Y \mid Z \vdash_{\mathsf{RI}} (X\multimap Y) \otimes Z}{
    \infer[\Pass 2 \LI]{X \multimap Y \mid Z \vdash_{\mathsf{LI}} (X\multimap Y) \otimes Z}{
      \infer[\F 2 \Pass]{X \multimap Y \mid Z \vdash_{\Pass} (X\multimap Y) \otimes Z}{
        \infer[\tr]{X \multimap Y \mid Z \vdash_{\F} (X\multimap Y) \otimes Z}{
          \infer[\lright]{X \lolli Y \mid \quad \vdash^{\bullet}_{\RI} X \lolli Y}{
            \infer[\LI 2 \RI]{X \lolli Y \mid X^{\bullet} \vdash^{\bullet}_{\RI} Y}{
              \infer[\Pass 2 \LI]{X \lolli Y \mid X^{\bullet} \vdash^{\bullet}_{\LI} Y}{
                \infer[\F 2 \Pass]{X \lolli Y \mid X^{\bullet} \vdash^{\bullet}_{\Pass} Y}{
                  \infer[\lleft]{X \lolli Y \mid X^{\bullet} \vdash^{\bullet}_{\F} Y}{
                    \infer[\LI 2 \RI]{- \mid X \vdash_{\RI} X}{
                      \infer[\Pass 2 \LI]{- \mid X \vdash_{\LI} X}{
                        \infer[\pass]{- \mid X \vdash_{\Pass} X}{
                          \infer[\Pass 2 \LI]{X \mid \quad \vdash_{\LI} X}{
                            \infer[\F 2 \Pass]{X \mid \quad \vdash_{\Pass} X}{
                              \infer[\ax]{X \mid \quad \vdash_{\F} X}{}
                            }
                          }
                        }
                      }
                    }
                    &
                    \infer[\Pass 2 \LI]{Y \mid \quad \vdash_{\LI} Y}{
                    \infer[\F 2 \Pass]{Y \mid \quad \vdash_{\Pass} Y}{
                    \infer[\ax]{Y \mid \quad \vdash_{F} Y}{}
                    }
                      }
                    }
                  }
                }
              }
            }
            &
                \infer[\LI 2 \RI]{- \mid Z \vdash_{\RI} Z}{
                  \infer[\Pass 2 \LI]{- \mid Z \vdash_{\LI} Z}{
                    \infer[\pass]{- \mid Z \vdash_{\Pass} Z}{
                      \infer[\Pass 2 \LI]{Z \mid \quad \vdash_{\LI} Z}{
                        \infer[\F 2 \Pass]{Z \mid \quad \vdash_{\Pass} Z}{
                          \infer[\ax]{Z \mid \quad \vdash_{\F} Z}{}
                        }
                      }
                    }
                  }
                }
          }
              }
            }
          }
\end{displaymath}
In the mean time, our focused system does not contain redundant non-determinacy.
For example, as the proof equivalences above, $\tr$ and $\lleft$ rule applications could be swapped when $\Gamma, \Delta$, and $\Delta'$ separated properly.
Without loose of generality, we assume $\Gamma, \Delta$, and $\Lambda$ are sufficient and necessary in proving each sequent respectively.
In tagged focused calculus, such exchangeable situation would not happen.
Given a sequent $A \lolli B \mid \Gamma , \Delta , \Lambda \vdash_{\RI} C \ot D$ and supposed that $B$ is irreducible, $C = A' \lolli B'$,
we can construct a proof:
\begin{displaymath}
  \infer[\LI 2 \RI]{A \lolli B \mid \Gamma , \Delta , \Lambda \vdash_{\RI} (A' \lolli B') \ot D}{
    \infer[\Pass 2 \LI]{A \lolli B \mid \Gamma , \Delta , \Lambda \vdash_{\LI} (A' \lolli B') \ot D}{
      \infer[\F 2 \Pass]{A \lolli B \mid \Gamma , \Delta , \Lambda \vdash_{\Pass} (A' \lolli B') \ot D}{
        \infer[\lleft]{A \lolli B \mid \Gamma , \Delta , \Lambda \vdash_{\F} (A' \lolli B') \ot D}{
          \deduce{- \mid \Gamma \vdash_{\RI} A }{\vdots}
          &
          \infer[\LI 2 \RI]{B \mid \Delta , \Lambda \vdash_{\RI} (A' \lolli B') \ot D}{
            \infer[\LI 2 \RI]{B \mid \Delta , \Lambda \vdash_{\LI} (A' \lolli B') \ot D}{
              \infer[\Pass 2 \LI]{B \mid \Delta , \Lambda \vdash_{\LI} (A' \lolli B') \ot D}{
                \infer[\F 2 \Pass]{B \mid \Delta , \Lambda \vdash_{\Pass} (A' \lolli B') \ot D}{
                  \infer[\tr]{B \mid \Delta , \Lambda \vdash_{\F} (A' \lolli B') \ot D}{
                    \deduce{B \mid \Delta \vdash^{\bullet}_{\RI} A' \lolli B'}{\vdots}
                    &
                    \deduce{- \mid \Lambda \vdash_{\RI} D}{\vdots}
                  }
                }
              }
            }
          }
        }
      }
    }
  }
\end{displaymath}
However, the other way around(decompose $\ot$ prior to $\lolli$) is impossible.
Because the sequent is black, if we want to decompose $A \lolli B$, we have to decompose $A' \lolli B'$ first to obtain $A^{\bullet}$ in context then send $\Gamma , \Delta , A'$ to left premise of $\lleft$ application.
In the end, the proof tree cannot close:
\begin{displaymath}
  \infer[\LI 2 \RI]{A \lolli B \mid \Gamma , \Delta , \Lambda \vdash_{\RI} (A' \lolli B') \ot D}{
    \infer[\Pass 2 \LI]{A \lolli B \mid \Gamma , \Delta , \Lambda \vdash_{\LI} (A' \lolli B') \ot D}{
      \infer[\F 2 \Pass]{A \lolli B \mid \Gamma , \Delta , \Lambda \vdash_{\Pass} (A' \lolli B') \ot D}{
        \infer[\tr]{A \lolli B \mid \Gamma , \Delta , \Lambda \vdash_{\F} (A' \lolli B') \ot D}{
          \infer{A \lolli B \mid \Gamma , \Delta \vdash^{\bullet}_{\RI} A' \lolli B'}{
            \infer[\lright]{A \lolli B \mid \Gamma , \Delta , A'^{\bullet} \vdash^{\bullet}_{\RI} B'}{
              \infer[\LI 2 \RI]{A \lolli B \mid \Gamma , \Delta , A'^{\bullet} \vdash^{\bullet}_{\RI} B}{
                \infer[\Pass 2 \LI]{A \lolli B \mid \Gamma , \Delta , A'^{\bullet} \vdash^{\bullet}_{\LI} B}{
                  \infer[\F 2 \Pass]{A \lolli B \mid \Gamma , \Delta , A'^{\bullet} \vdash^{\bullet}_{\Pass} B}{
                    \infer[\lleft]{A \lolli B \mid \Gamma , \Delta , A'^{\bullet} \vdash^{\bullet}_{\F} B}{
                      \deduce{- \mid \Gamma , \Delta , A'^{\bullet} \vdash_{\RI}}{??}
                      &
                      \deduce{B \mid \quad \vdash_{\LI} B'}{??}
                    }
                  }
                }
              }
            }
          }
          &
          \deduce{- \mid \Lambda \vdash_{\RI} D}{\vdots}
        }
      }
    }
  }
\end{displaymath}
The situation is worse if $C$ is positive, we cannot produce tags from $\lright$ application to ensure applicability of $\lleft$, so the whole derivation gets stuck and cannont close.
This fact shows that in tagged focused system, we cannot permute $\tr$ and $\lleft$ applications arbitrarily.
In this sense, two equivalent proofs in SkNMILL become identical in focused calculus.
Therefore, tagged focused system here is not only sound but also without redundant non-determinacy.

Moreover, not only $\tr$ and $\lleft$ equivalent case becomes identical in focused system, but any other equivalent cases hold the same fact.
We can substantiate statement above by following results.
\begin{lemma}\label{AdmissibleInRI}
  Following rules are admissible in tagged focused system:
  \begin{displaymath}
    \infer[\unitl^{\RI}]{\I \mid \Gamma \vdash_{\RI} C}{- \mid \Gamma \vdash_{\RI} C}
    \quad
    \infer[\tl^{\RI}]{A \mid B, \Gamma \vdash_{\RI} C}{A \mid B, \Gamma \vdash_{\RI} C}
    \quad
    \infer[\pass^{\RI}]{- \mid \Gamma \vdash_{\RI} C}{A \mid \Gamma \vdash_{\RI} C}
  \end{displaymath}
  \begin{displaymath}
    \infer[\lleft^{\RI}]{A \lolli B \mid \Gamma , \Delta \vdash_{\RI} C}{
    - \mid \Gamma \vdash_{\RI} A
    &
    B \mid \Delta \vdash_{\RI} C
    }
    \quad
    \infer[\ax^{\RI}]{A \mid \quad \vdash_{\RI} A}{}
    \quad
    \infer[\unitr^{\RI}]{- \mid \quad \vdash_{\RI} \I}{}
  \end{displaymath}
  \begin{displaymath}
    \infer[\tr^{\RI}]{S \mid \Gamma , \Delta \vdash_{\RI} \ldbc \Gamma' \mid A \rdbc_{\lolli} \ot B}{
      S \mid \Gamma , \Gamma' \vdash_{\RI} A
      &
      - \mid \Delta \vdash_{\RI} B
    }
  \end{displaymath}
\end{lemma}

$\tr^{\RI}$ is a generalized version of $\tr$.
We endorse it in order to strengthen inductive hypothesis and make our proof smoothly.
For example, considering following proof tree, if we invoke simple version of $\tr$ in phase $\RI$, then we cannot proof its admissibility:
\begin{displaymath}
  \infer[\tr^{\RI}_s]{S \mid \Gamma , \Delta \vdash_{\RI} (A' \lolli B') \ot B}{
    \infer[\lright]{S \mid \Gamma \vdash_{\RI} A' \lolli B'}{
      \deduce{S \mid \Gamma , A' \vdash_{\RI} B'}{f}
    }
    &
    \deduce{- \mid \Delta \vdash_{\RI} B}{g}
  }
\end{displaymath}
In this case, we cannot permute $\tr^{\RI}_s$ up to use inductive hypothesis.
Therefore we need to have a stronger version of $\tr$ which performs $\tr$ and $\lright$ in the same time.
There is no generality loss because we can always take $\Gamma'$ as empty then reduce to simple $\tr$.

If we have a derivation in focused system, then we can define a function $\mathsf{emb}$ which erases phases and tags to obtain a derivation in original unfocused seqeuent calculus.
On the other hand, we can define a function $\mathsf{focus}$ to map each derivation in cut-free sequent calculus to a derivaiton in focused calculus because lemma \ref{AdmissibleInRI} holds.
Specifically, for each derivation in cut-free sequent calculus, function $\mathsf{focus}$ adds phase label $\RI$ on $\vdash$ and rename each rule into its $\RI$ version.

Moreover, we can show that two equivalent derivations in SkNMILL become identical in focused seqeunt calculus.
Therefore, deciding two proofs to be equivalent could be done by checking whether their corresponding focused derivation are identical.
%Next, we can extend this result to decide equality of morphisms in skew monoidal closed categories.

We take one more step to extend the results to categorical side.
According to categorical semantics section above, we know SkNMILL is sound and complete with free skew monoidal categories, so we can construct an effective procedure to determine equivalence of morphisms in a free skew monoidal closed category.
\begin{enumerate}
  \item Given two moprhisms $f$ and $g$ in a free skew monoidal closed category.
  \item Use $\mathsf{complt}$ function to obtain two derivations $\mathsf{complt} \text{ } f$ and $\mathsf{complt} \text{ } g$ in sequent calculus system.
  \item Apply $\mathsf{complt} \text{ } f$ and $\mathsf{complt} \text{ } g$ to function $\mathsf{focus}$ respectively.
  \item If $\mathsf{focus} \text{ } (\mathsf{complt} \text{ } f) = \mathsf{focus} \text{ } (\mathsf{complt} \text{ } g)$, then $f \doteq g$, else $f \not\doteq g$.
\end{enumerate}

%%\begin{theorem}
%%  \begin{enumerate}
%%    \item For any derivation $f : S \mid \Gamma \vdash_{\mathsf{RI}} C$, there is a derivation $\mathsf{emb}$ $f : S \mid \Gamma \vdash C$.
%%    \item For any derivation $f : S \mid \Gamma \vdash C$, there is a derivation $\mathsf{focus}$ $f : S \mid \Gamma \vdash_{\mathsf{RI}} C$.
%%  \end{enumerate}
%%\end{theorem}
%%\begin{theorem}
%%  \begin{enumerate}
%%    \item For any $f : S \mid \Gamma \vdash_{\RI} C$, $\mathsf{focus} \text{ } (\mathsf{emb} \text{ } f) = f$.
%%    \item For any $f, g : S \mid \Gamma \vdash C$,  if $f \circeq g$, then $\mathsf{focus} \text{ } f = \mathsf{focus} \text{ } g$.
%%    \item For any $f : S \mid \Gamma \vdash C$, $\mathsf{emb} \text{ } (\mathsf{focus} \text{ } f) \circeq f$.
%%  \end{enumerate}
%%\end{theorem}

\section{Conclusion}

  \bibliographystyle{eptcs}
  \bibliography{docRefs}
\end{document}
